---
title: "Data Exploration Part 1"
output: html_notebook
---


# Business Question 1: Does the weather effectively influence the flights departure delays?

```{r}
library(tidyverse)
library(GGally)
```




```{r}
#flights_weather <- read_csv(here::here("cleaned_data/flights_weather.csv"))
tight_flights_weather <- read_csv(here::here("cleaned_data/tight_flights_weather.csv"))
```



```{r}
skimr::skim(tight_flights_weather)
```




```{r}
tight_flights_weather %>% 
  filter(!is.na(arr_delay)) %>% 
  summarise(cor(dep_delay, arr_delay))
```





```{r}
tight_flights_weather %>% 
  select(dep_delay, wind_dir, wind_speed, wind_gust, visib) %>% 
  ggcorr(label = TRUE)
```


At a first sight, looking at the correlation indexes, it seems that there is not so much interaction between `dep_delays` and `weather` conditions.

I also notice that there is a very strong correlation between `wind speed` and `wind gust speed` for this reason, I could consider to delete one of the two since their relationship is so strong.



```{r}
tight_flights_weather %>% 
  select(wind_dir, wind_speed, wind_gust, visib, dep_delay) %>% 
  ggpairs(progress = FALSE)
```

It seems interesting to analyse the relationship that `dep_delays` has with `wind_dir` and `visibility`.

```{r}
tight_flights_weather %>%
  group_by(month) %>% 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = month, y = mean_dep_delay)+
  geom_col(fill = "9900FF")+
  scale_x_continuous(n.breaks = 12)+
  labs(
    x = "\nmonths",
    y = "departure delays mean (in mins)\n",
    title = "\nDeparture delays per month in 2017\n"
  )+
  theme_light()
```

# Wind Speed

```{r}
tight_flights_weather %>% 
  ggplot()+
  aes(x = wind_speed, y = dep_delay)+
  geom_point()
```

From this plot it seems that the `wind_speed` is more related maybe to the number of delays than to their length, but I am not actually interested in that, since I decided to concentrate on their length. 

```{r}
tight_flights_weather %>% 
  ggplot()+
  aes(x = wind_dir, y = dep_delay)+
  geom_point()
```


 
 
```{r}
tight_flights_weather %>%
  group_by(month, wind_speed_group) %>% 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = month, y = mean_dep_delay, fill = wind_speed_group)+
  geom_col(position = "dodge")+
  #geom_line() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  theme_light()+
  labs(
    x = "\n months",
    y = "departure delays mean (in mins)\n",
    title = "\nDeparture delays depending on different wind speed conditions\n"
  )+
  scale_x_continuous(n.breaks = 12)
```
```{r}
tight_flights_weather %>%
  group_by(month, wind_speed_group) %>% 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = month, y = mean_dep_delay, colour = wind_speed_group)+
  #geom_col(position = "dodge")+
  geom_line()+
  geom_point()
```



Generally speaking the wind speed is not regular during the year even if, having a look of the below plot, it seems that the departure delays seems to be more uniformly distributed with a few exceptions during winter. This would suggest that there is not a great relationship between departure delays and wind speed, as we already noticed looking at the correlation coefficients.

However, looking with more attention at the above plot, we can notice that, at least, month by month it seems that the length of the delays increase when the wind speed is particularly high (look at May) even if even this proportion doesn't seems to be regular. 



```{r}
tight_flights_weather %>% 
  group_by(month) %>% 
  summarise(mean_wind_speed = mean(wind_speed, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = month, y = mean_wind_speed)+
  geom_col(fill = "339995")+
  scale_x_continuous(n.breaks = 12)+
  labs(
    x = "\nmonths",
    y = "mean of wind speed per month 2017\n",
    title = "\nMean of wind spped per month in 2017\n"
  )+
  theme_light()
```



```{r}
tight_flights_weather %>% 
  ggplot()+
  aes(x = wind_speed_group, y = dep_delay)+
  geom_boxplot()
```

```{r}

model <- lm(dep_delay ~ wind_speed_group, tight_flights_weather)
summary(model)
```




## Visib





```{r}
tight_flights_weather %>% 
  group_by(month, visib_group) %>% 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x = month, y = mean_dep_delay, fill = visib_group)+
  geom_col(position = "dodge")+
  #geom_line() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) + 
  theme_light()+
  labs(
    x = "\n months",
    y = "departure delays mean (in mins)\n",
    title = "\nDeparture delays depending on different wind speed conditions\n"
  )+
  scale_x_continuous(n.breaks = 12)
```

It seems that the visibility can be considered as a better parameter for measuring the departure delays than the wind speed.


However, since we saw that the correlation coefficient is still very low, we can deduce that even if its true that in mean less visibility is in some way proportional to the length of delays it cannot be considered as a significant parameter.



```{r}
tight_flights_weather %>% 
  ggplot()+
  aes(x = visib_group, y = dep_delay)+
  geom_boxplot()
```


```{r}
tight_flights_weather %>% 
  distinct(wind_dir)
```



```{r}
tight_flights_weather %>% 
  ggplot()+
  aes(x = wind_dir, y = dep_delay)+
  geom_point()
```


I don't find from this plot a particular incidence of `wind_dir` with the `dep_delay`.































































































